---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

Load the data downloaded from the GitHub repo

1. Load the data (i.e. read.csv())
```{r}
df <- read.csv("activity.csv", head=TRUE)
#summary(df)
```

2. Process/transform the data (if necessary) into a format suitable for your analysis
```{r}
library(dplyr)
by_day <- group_by(df, date)
day_step_by_int <- summarize(by_day, total_int = sum(steps, na.rm=T))
```

## What is mean total number of steps taken per day?

1. Calculate the total number of steps taken per day 

```{r}
b <- day_step_by_int[,2]
b
```


```{r}
hist(b$total_int, 10)
my_mean <- mean(b$total_int)
my_median <- median(b$total_int)
```

The mean is `r my_mean` and the median is `r my_median`.

## What is the average daily activity pattern?

```{r}
by_interval <- group_by(df, interval)
my_mean <- summarize(by_interval, mean(steps, na.rm=T))
colnames(my_mean) <- c("Interval", "Mean.Steps")
plot(my_mean$Interval, my_mean$Mean.Steps, type="l", main="Average Daily Activity Pattern", xlab="Interval", ylab="Average Steps")
```

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
