---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r setoptions, echo=TRUE}
require(knitr)
opts_chunk$set(echo = TRUE, cache = TRUE, cache.path = "cache/", fig.path="figure/")
```

## Loading and preprocessing the data

###Load the data downloaded from the GitHub repo

####Step 1. Load the data (i.e. `read.csv()`)
```{r}
df <- read.csv("activity.csv", head=TRUE)
#summary(df)
```

####Step 2. Process/transform the data into a suitable analytical format 
```{r}
library(dplyr)
by_day <- group_by(df, date)
#day_step_by_int <- summarize(by_day, total_int = sum(steps, na.rm=T))
```

## What is mean total number of steps taken per day?

###Analyze the data as it exists

####Step 1. Calculate the total number of steps taken per day 

```{r}
day_step_by_int <- summarize(by_day, total_int = sum(steps, na.rm=T))
```

####Step 2. Make a histogram of the total number of steps taken each day

```{r part2_2}
hist(day_step_by_int$total_int, 10, main="Histogram of Frequency of Daily Steps", 
     xlab="Number of Daily Steps", ylab="Frequency")
my_mean <- mean(day_step_by_int$total_int)
my_median <- median(day_step_by_int$total_int)
```

####Step 3. Calculate and report the mean and median of the total number of steps taken per day

Response: The mean is `r my_mean` and the median is `r my_median`.

## What is the average daily activity pattern?

###Analyze the data as it exists

####Step 1. Make a time series plot (i.e. `type = "l"`) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r part3_1}
by_interval <- group_by(df, interval)
int_mean <- summarize(by_interval, mean(steps, na.rm=T))
colnames(int_mean) <- c("Interval", "Mean.Steps")
plot(int_mean$Interval, int_mean$Mean.Steps, type="l", 
     main="Average Daily Activity Pattern", xlab="Interval", ylab="Average Steps")
```

####Step 2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
max_interval <- int_mean$Interval[which.max(int_mean$Mean.Steps)]
```

Response: The interval with the maximum average steps across the entire day is `r max_interval`

## Imputing missing values

###Impute missing values and analyze the resulting data

####Step 1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r}
missing_values <- sum(is.na(by_day$steps))
```

Response: The total number of missing values in the dataset is `r missing_values`

####Step 2. Devise a strategy for filling in all of the missing values in the dataset. 

Response: The relatively unsophisiticated strategy chosen to impute the missing values is to calculate the mean of each 5-minute interval and subsititue that for the `NA` values.

```{r}
int_mean <- mutate(int_mean, Step.Missing = round(Mean.Steps))
```

####Step 3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
adj_interval <- merge(by_interval, int_mean, by.x = "interval", by.y = "Interval")
adj_interval <- mutate(adj_interval, 
                       imputted = ifelse(is.na(adj_interval$steps), 
                                         adj_interval$Step.Missing, 
                                         adj_interval$steps))
adj_day <- group_by(adj_interval, date)
adj_step_by_day <- summarize(adj_day, adj_steps = sum(imputted))
```

####Step 4a. Make a histogram of the total number of steps taken each day.

```{r part4_4a}
hist(adj_step_by_day$adj_steps, 10, main="Histogram of Frequency of Imputed Daily Steps", 
     xlab="Number of Daily Steps", ylab="Frequency")

```

####Step 4b. Calculate the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

```{r}

original_mean <- mean(day_step_by_int$total_int)
original_median <- median(day_step_by_int$total_int)

imputed_mean <- mean(adj_step_by_day$adj_steps)
imputed_median <- median(adj_step_by_day$adj_steps)

diff_mean = imputed_mean - original_mean
diff_median = imputed_median - original_median

```

Response: The mean for the data with imputed values is `r imputed_mean` while the
median is `r imputed_median`. This compares to the origial values of `r original_mean`
and `r original_median` respectively.

By introducing the imputed values the mean increased by `r diff_mean` and the median
increased by `r diff_median`. By using the imputed values we reduced the number
of days that had very few steps taken and increased the number of days that
represent the mean. The number of days in the historgram above the mean did not
change.

## Are there differences in activity patterns between weekdays and weekends?

###Determine if the day of the week makes a difference in the number of steps in the imputed data

####Step 1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

####Step 2. Make a panel plot containing a time series plot (i.e. `type = "l"`) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 


